---
title: '"Getting IUCN species for tet region"'
output: html_document
---

#Summary 

This script compares the shapefile of Tetiaroa and its surrounding waters to the species range map shapefiles from the IUCN to extract all the species in this area. Methods match `1_get_ne_iucn_spp.Rmd` by Jamie Montgomery, OHI Northeast, and also use data and code developed by Casey O'Hara: https://github.com/madelinemberger/spp_risk_dists.

#Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

library(sf)
library(raster)
library(rfishbase) #included in ne, may not need
library(rredlist)
library(taxize)
library(here)
library(tidyverse)

source(here('workflow/R/common.R'))

#tetiaroa region shape file
tet_shp <- tet_buffer_3nm


# api keys for IUCN 

api_file <- file.path(dir_M, 'git-annex/globalprep/spp', 
                      'api_key_gc.csv')
api_key <- scan(api_file, what = 'character')

```

#Methods

First step is to get a list of the cell IDs contained in the tet_region

```{r}

#these are global cell ids in raster format pulled from the spp_risk_dists repo.
cells <- raster("~/github/spp_risk_dists/_spatial/cell_id_rast.tif") #resolution is 10 km - is this too big? the area of this shp is 259 km2

#reproject shapefile to the same crs as the raster, which is WGS84

tet_reproj <- st_transform(tet_shp, crs = "+proj=cea +lon_0=0 +lat_ts=45 +x_0=0 +y_0=0 +ellps=WGS84
+units=m +no_defs")

plot(cells)
plot(tet_reproj, add = T)

#use extract to get the region ids for just the Tetiaroa area

tet_cells <- raster::extract(cells, tet_reproj) %>% 
  unlist() #this creates a vector out of the list from extract

```

Read in files that contain all the species ranges and information

```{r}

# .csv file that lists all species and the file path to their map
spp_maps <- read_csv('~/github/spp_risk_dists/_data/spp_marine_maps_2019-2.csv',
                     col_types = 'ddciccc')

#file with species information to link to iucn_sid at the end
spp_info <- read_csv("/home/ohara/git-annex/spp_risk_dists/iucn/spp_info_from_api_2019-2.csv")

```

The following forloop goes through each species:  
- finds it's species path
- reads in the species .csv that lists all global cells where it is found
- filters that list to just those in the Northeast
- returns an aggregated dataframe (`taxa_cells_df`) with all species scientific names, their unique ids (sids), and cellIDs

```{r}

#grab each taxa's folder filepath
taxa <- spp_maps$dbf_file %>% 
  unique() %>% 
  str_replace('\\....$', '')

#create an empty list that is the length of all taxa. We are going to fill this list  
taxa_cells_list <- vector('list', length = length(taxa))

#for each taxa, grab the species map (raster) and filter to only keep those cells in the northeast.
for(i in seq_along(taxa)) { ### i <- 5
    taxon <- taxa[i]
    print(i)
    spp_ids_in_taxon <- spp_maps %>%
      filter(str_detect(dbf_file, taxon)) %>%
      .$iucn_sid
    cat(sprintf('processing %s spp in %s...\n', length(spp_ids_in_taxon), taxon)) #help with this, message?
    
    spp_cells <- parallel::mclapply(spp_ids_in_taxon, mc.cores = 32,
                                    FUN = function(x) { ### x <- spp_ids_in_taxon[1]
                                      f <- file.path('/home/ohara/git-annex/spp_risk_dists/spp_rasters_2019',
                                                     sprintf('iucn_sid_%s.csv', x))
                                      if(file.exists(f)) {
                                        y <- read_csv(f, col_types = 'di') %>%
                                          mutate(iucn_sid = x) %>%
                                          select(-presence)  %>%
                                          filter(cell_id %in% tet_cells)
                                      } else {
                                        y <- data.frame(cell_id = NA,
                                                        iucn_sid = x, 
                                                        f = f, error = 'file not found')
                                      }
                                      return(y)
                                    }) %>%
      bind_rows() %>%
      mutate(spp_gp = taxon)
    
    taxa_cells_list[[i]] <- spp_cells
}

taxa_cells_list[[3]] #was using to just look at a few
taxa_cells_df <- taxa_cells_list %>%
    bind_rows() %>% 
    filter(!is.na(cell_id)) %>% 
    select(iucn_sid) %>% 
    distinct() %>%
    left_join(spp_info) #these all came out Na, so this df is empty 
```

Small note on looking at species maps - use filepath?
```{r}

humpback_whale <- spp_maps %>% 
  filter(iucn_sid == 13006)

```

Attach common names

```{r}
#create list of scintific names of species

scinames <- taxa_cells_df$sciname

#create loop around redlist function rl_search to find common names  
#create empty data frame that will be filled
iucn_common_names <- data.frame() 

for(i in 1:length(scinames)){
  sp <- scinames[i]
  print(i)
  if(sp %in% c("Procelsterna cerulea")){ #this one threw some error and no $result was returned from rl_search
  comm <- NA
  }else{
  comm <- rredlist::rl_search(sp, key = api_key )$result$main_common_name
  }
  
  df <- data.frame(sciname = sp,
                   common = comm)
  
  iucn_common_names <- rbind(iucn_common_names, df) %>%
    mutate(common = as.character(common),
         sciname = as.character(sciname))
}

#had to remove blue noddy because it was causing an error

```

Check out the ones we are still missing for common name

```{r}

miss_sp <- iucn_common_names %>%
  filter(is.na(common))

```

We are still missing `r nrow(filter(iucn_common_names, is.na(common)))`. Use taxize package to try and find these. (note, above is showing up as 8 but there are 17 missing?). 

```{r}

#create vector from missing scinames
missing_scinames <- iucn_common_names%>%
  filter(is.na(common)) %>%
  .$sciname

#loop to pull out the common names from taxize

common_names <- c()
for(i in 1:length(missing_scinames)){ #i <- 1
  sp <- missing_scinames[i]
  comm <- taxize::sci2comm(sp, db = "itis")
  print(comm)
 
  common_names <- c(common_names, comm)


}

#output is list, turn into dataframe

df <- do.call(rbind,lapply(common_names,data.frame))
df$sciname <- rownames(df)

```

This only got us blue noddy, not really helpful. Try some exploration

```{r}
#this is a sea cucumber, does not have a common name in IUCN
rl_search("Holothuria olivacea", key = api_key )$result$main_common_name

taxize::get_wormsid("Holothuria olivacea")

scinames_explore <- miss_sp %>%
  separate(sciname, sep = " ", into = c("genus", "family")) %>% 
  group_by(genus) %>% 
  summarize(
    count = length(family)
  )


```

There are only 6 genus' represented in the missing data: 
- 8 of them are Holothuria, which are sea cucumbers
- `scinames_explore$genus[1]` are slipper lobsters
- `scinames_explore$genus[3]` are also sea cucumbers
- `scinames_explore$genus[4]` mostly deep sea blind lobsters
- `scinames_explore$genus[5]` these are the noddy's will add in manually below
- `scinames_explore$genus[6]` more lobsters?

```{r}

spp_names <- iucn_common_names %>% 
  mutate(common_name = case_when(
              sciname == "Procelsterna cerulea" ~ "Blue Noddy",
              TRUE ~ as.character(common)))

#add in the rest of the info 

iucn_spp_info <- taxa_cells_df %>%
  select(iucn_sid, sciname, population, category) %>% 
  left_join(., spp_names, by = "sciname") %>% 
  select(-common) %>% 
  distinct()

```

Write out csv to folder `int`

```{r}

#save
write_csv(iucn_spp_info, "int/1_iucn_spp_in_tet.csv")

```

Last step is to get a list of the maps now that we have the names before moving onto to script 2
(come back to this since the subpop region is tricky with this region)

Note: the marine maps df only has 128 rows, meaning we are missing 5 maps
```{r}
# only loggerhead sea turtle has subpops so going to filter out the one we want
spp_marine_maps <- read_csv("~/github/spp_risk_dists/_data/spp_marine_maps_2018-1.csv") %>%
  filter(sciname %in% iucn_spp_info$sciname, 
         subpop %in% c("Pacific South", NA))

write.csv(spp_marine_maps, "int/1_iucn_spp_shp_filepaths.csv")

```

##Adding coral 

The above process for some reason did not have coral. It also may have been missing some of the species we had in ICO, and we want all those to be on this list as well. This section will work on adding these missing groups to the spp list. 

Note: also check IUCN id, noticed that they might be different last time. 

1. Get files needed 
```{r}

#read in ico species list

ico_sp <- read_csv(here("/prep/sp/ico/v2020/int", "tet_ico_all.csv"))

spp_sp <- read_csv(here("/prep/bd/spp/v2020/int","1_iucn_spp_in_tet.csv")) %>% 
  mutate(
    sciname = str_to_lower(sciname),
    name = str_to_lower(common_name)
  ) %>% 
  dplyr::select(-common_name, name, sciname, iucn_sid, population, category)

```

We can detect any differences using `setdiff()`?

```{r}
#this is a list of whats in ico but NOT in spp - this is what we need to add

ico_not_sp <- setdiff(ico_sp$sciname, spp_sp$sciname) 

#this is a list of whats in spp but NOT ICO, just as a reference
setdiff(spp_sp$sciname, ico_sp$sciname) 

```

Filter the ico list for the missing ones and bind to the iucn list

```{r}

iucn_all <- ico_sp %>% 
  filter(sciname %in% ico_not_sp) %>% 
  bind_rows(., spp_sp) #create new dataframe with all the species


```

This got us 4 species of coral: 

- lobe coral	porites lobata
- staghorn coral	acropora striata	
- knob-horned coral	pocillopora verrucosa	
- fire coral	millepora platyphyll

According to [this](https://aslopubs.onlinelibrary.wiley.com/doi/full/10.1002/lno.10231() 2013 paper, there should also be (at least) a few species in the following genus':

- pavora
- astrea
- montipora
- leptastrea

We'll try using the French Polynesia region to find more corals. 
```{r}

#"porites", "acropora", "millepora", "pocillopora"

genus <- c("pavora", "astrea", "montipora", "leptastrea") #when you add more though it doesn't quite work

coral_fp <- rl_sp_country('PF', key = api_key)$result %>%
  mutate(country_code = "PF") %>%
  rename("iucn_sid" = "taxonid") %>% 
  mutate(
    sciname = str_to_lower(scientific_name)
  ) %>%
  filter(str_detect(sciname, pattern = genus))

#got montipora and leptastrea  

```

The list above is helpful but its hard to know if all those corals are found on Tetiaroa. Also try map extraction method to be safe: 

1. Get a list of all the cell ids in the FP region

```{r}
#get FP region, rgn_id = 147

fp <- regions_shape() %>% 
  filter(rgn_id == 147) %>% 
  filter(rgn_type == "eez") %>% 
  st_transform(., crs = "+proj=cea +lon_0=0 +lat_ts=45 +x_0=0 +y_0=0 +ellps=WGS84
+units=m +no_defs")

#these are global cell ids in raster format pulled from the spp_risk_dists repo.
cells <- raster("~/github/spp_risk_dists/_spatial/cell_id_rast.tif") #resolution is 10 km - is this too big? the area of this shp is 259 km2

#reproject shapefile to the same crs as the raster, which is WGS84

plot(cells)
plot(fp, add = T)

#use extract to get the region ids for just the Tetiaroa area

fp_cells <- raster::extract(cells, fp) %>% 
  unlist() #this creates a vector out of the list from extract


```

For loop for pulling out species:

```{r}

#this isn't running 
#grab each taxa's folder filepath
taxa <- spp_maps$dbf_file %>% 
  unique() %>% 
  str_replace('\\....$', '')

#create an empty list that is the length of all taxa. We are going to fill this list  
taxa_cells_list <- vector('list', length = length(taxa))

#for each taxa, grab the species map (raster) and filter to only keep those cells in the northeast.
for(i in seq_along(taxa)) { ### i <- 5
    taxon <- taxa[i]
    print(i)
    spp_ids_in_taxon <- spp_maps %>%
      filter(str_detect(dbf_file, taxon)) %>%
      .$iucn_sid
    cat(sprintf('processing %s spp in %s...\n', length(spp_ids_in_taxon), taxon)) #help with this, message?
    
    spp_cells <- parallel::mclapply(spp_ids_in_taxon, mc.cores = 32,
                                    FUN = function(x) { ### x <- spp_ids_in_taxon[1]
                                      f <- file.path('/home/ohara/git-annex/spp_risk_dists/spp_rasters_2019',
                                                     sprintf('iucn_sid_%s.csv', x))
                                      if(file.exists(f)) {
                                        y <- read_csv(f, col_types = 'di') %>%
                                          mutate(iucn_sid = x) %>%
                                          select(-presence)  %>%
                                          filter(cell_id %in% fp_cells)
                                      } else {
                                        y <- data.frame(cell_id = NA,
                                                        iucn_sid = x, 
                                                        f = f, error = 'file not found')
                                      }
                                      return(y)
                                    }) %>%
      bind_rows() %>%
      mutate(spp_gp = taxon)
    
    taxa_cells_list[[i]] <- spp_cells
}
```

