---
title: "Species status and trend"
output: html_document
---

#Summary 

This script joins the different conservation status to their respective weights, in addition to creating a trends layer using the `rl_search()` function in the rredlist package

#Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

library(rredlist)
library(here)
library(tidyverse)
source(here('workflow/R/common.R'))

#file paths
goal     <- 'spp'
scenario <- 'v2020'
dir_anx       <- file.path(dir_M, 'git-annex/globalprep') 
dir_goal      <- file.path(here('prep/bd',goal, scenario))
dir_goal_anx  <- file.path(dir_anx, goal, scenario)
dir_github    <- '~/github/tet-prep'


# api keys for IUCN 

api_file <- file.path(dir_M, 'git-annex/globalprep/spp', 
                      'api_key_gc.csv')
api_key <- scan(api_file, what = 'character')

```

#Status

Read in species list

```{r}

spp_list <- read_csv(here("/prep/bd/spp/v2020/int", "2_spp_ico_iucn_50.csv")) %>%
      dplyr::select(name, sciname, iucn_sid, category) %>%
      distinct()


```

First, just score this list

```{r}

pop_cat <- data.frame(category = c("LC", "NT", "VU", "EN", "CR", "EX", "T", "CD", "NE", "DD"), 
                      risk_score = c(   0,  0.2,  0.4,  0.6,  0.8,  1.0, 0.6,  0.3,   NA,  NA),
                      stringsAsFactors = FALSE) %>% 
  mutate(
    stat_score = 1 - risk_score
  )

#join the scores data to the list data - note, this doesn't have year. 

spp_status <- left_join(spp_list, pop_cat, by = "category") %>%
  mutate(stat_score = case_when(
    #category %in% c("LR/lc") ~ 1.0, #this one isn't showing up? double check 
   # category %in% c("LR/nt") ~ 0.8,
    category %in% c("LR/cd") ~ 0.9,
    TRUE ~ stat_score)) %>% 
  dplyr::select(-risk_score)

#double check that this is what the layer should look like 
```

#Trend
Next, get trends using the method Jamie does
```{r}

#forloop for each species to grab trend

df_iucn <- data.frame()
for(i in 1:nrow(spp_list)){
  
  print(i)
  sp <- as.character(spp_list[i,2]) #grab scientific name
  
  tr <- rl_search(sp, key = api_key)$result$population_trend
  
  if(is.null(tr)){
  
  df2 <- data.frame(sciname = sp,
                    trend = NA)
  }else{
    df2 <- data.frame(sciname = sp,
                      trend = tr)
  }
  
  df_iucn <- rbind(df_iucn, df2)
}

#write out to int folder

write.csv(df_iucn, "3_iucn_population_trends.csv")
```

##Score
Now, score the trend based on what was pulled from rl_search
```{r}

iucn_trends <- read_csv(here("prep/bd/spp/v2020/int", "3_iucn_population_trends.csv"))

#score table
pop_trend_score <- data.frame(trend = c("Increasing", "Decreasing", "Stable", "Unknown"),
                              score = c(0.025, -0.025, 0, NA))
```

Join to the trend list

```{r}

spp_trend_scores <- iucn_trends %>% 
  left_join(pop_trend_score) %>%
  filter(!is.na(trend)) #filter out three where trend was unknown
```

##Visualize

```{r}
spp_trends_plot <- spp_trend_scores %>%
  group_by(trend) %>%
  summarize(count = n())

spp_trends_plot

ggplot(spp_trends_plot, aes(x = trend, y = count)) +
  geom_bar(stat = "identity") +
  theme_bw()

#over half of them are unknown
```

##Save
```{r}

spp_trends <- spp_trend_scores %>%
  mutate(year = 2020) %>%   #have to add a year for the toolbox - should this be 2020
  dplyr::select(sciname, score, year)

write.csv(spp_trends, file = "~/github/tet-scores/region/layers/spp_trend.csv", row.names = F)  #save to toolbox


```

