---
title: "Tetiaroa RRedlist"
author: "Maddie Berger"
date: "6/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rredlist)
library(tidyverse)
library(dplyr)
library(raster)
library(here)
library(janitor)
source(here('workflow/R/common.R'))

api_file <- file.path(dir_M, 'git-annex/globalprep/spp', 
                      'api_key_gc.csv')
api_key <- scan(api_file, what = 'character')
```

Get US Minor French Polynesia species 
```{r}
spp_country_url <- 'http://apiv3.iucnredlist.org/api/v3/country/getspecies/id/%s?token=%s'
# 
# /api/v3/country/getspecies/:country?token='YOUR TOKEN'
countries <- rl_countries(key = api_key)$results
#UM is United States Minor Outlying Islands country code
# UM species risks
spp_um <- rl_sp_country('UM', key = api_key)$result %>%
  mutate(country_code = "UM") %>%
  rename("iucn_sid" = "taxonid")
##PF is French Polynesia country code
## French Polynesia Species risks
spp_fp <- rl_sp_country('PF', key = api_key)$result %>%
  mutate(country_code = "PF") %>%
  rename("iucn_sid" = "taxonid") %>% 
  mutate(
    scientific_name = str_to_lower(scientific_name)
  )

```

Filter out species we want for Tetiaroa:
```{r}
#read in csv with list of tet species

tet_species <- read_csv(here("./prep/bd/spp/v2020/raw_data", "tet_species.csv"))%>%
  clean_names() %>% 
  mutate(
    scientific_name = str_to_lower(scientific_name)
  )

tet_species_iucn <- spp_fp %>% 
  filter(scientific_name %in% tet_species$scientific_name)

tet_species_iucn <- inner_join(tet_species_iucn,tet_species, by = "scientific_name") %>%  dplyr::select(iucn_sid, scientific_name,category.x,name,ico)

```

Check to see if they were all there:
```{r}

missing_spp <- setdiff(tet_species$scientific_name, tet_species_iucn$scientific_name)

tet_species_missing_fp <- tet_species %>% 
  filter(scientific_name %in% missing_spp)
```

Seems like the French Polynesia list is missing some species that are found on Tetiaroa
https://rdrr.io/cran/rredlist/man/rl_occ_country.html

```{r}

#test hawksbill turtle, which is critically endangered

rl_occ_country("Eretmochelys imbricata", key = api_key) 
#can see that FP isn't on here, even though the range maps shows it occurs there


```

Solution is to filter out these missing species from the overall dataset
```{r}
## get a dataframe of all iucn redlist species

out <- rl_sp(all = TRUE, key = api_key)

all_df <- do.call(rbind, lapply(out, "[[", "result")) #creates data frames from each item in the large list in `out` and then binds them

all_df_comp <- all_df %>%
  dplyr::select(-infra_rank, -infra_name) %>%
    dplyr::rename(iucn_sid = taxonid, sciname = scientific_name) %>% 
    setNames(names(.) %>%
               stringr::str_replace('_name', '')) %>% 
  mutate(
    sciname = str_to_lower(sciname)
  )


#filter large df by the missing_spp vector

tet_not_listed <- all_df_comp %>%
  filter(sciname %in% missing_spp)

#this only has 7 obs, so I think we are still missing some

not_in_iunc <- setdiff(tet_species_missing_fp$scientific_name, tet_not_listed$sciname) #49 in this list

no_record <- tet_species %>% 
  dplyr::filter(scientific_name %in% not_in_iunc) 

#this seems really unlikely. something might be up with spelling?


```

So far, have created the following dfs:
- `tet_species_iucn` is all the species that matched what was pulled from the IUCN French Polynesia list
- `tet_species_missing_fp` is all the species that were missing from the IUCN French Polynesia list but were included in Tetiaroa's species list
- `tet_not_listed` is all the species that were matched using the entire IUCN dataset, but here are only 7 additional ones here
- `no_record` includes all the species we still can't match. This list is probably a result of differeces in spelling as there are many species on there that should be in the IUCN, like humpback whale, spotted eagle ray, and olive ridley's turtle. 

Join tet_species_iucn and tet_not_listed
```{r}

tet_species_clean <- tet_species_iucn %>% 
  dplyr::select(iucn_sid, scientific_name = sciname, category.x= category)

```

