---
title: 'OHI Tetiaroa: Fishing Regulations Resilience Layer'
author: 'Madeline Berger'
output:
  html_document:
    highlight: haddock
    includes:
      in_header: '~/github/tet-prep/workflow/templates/tet_hdr.html' 
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
---
# Summary 

Regulatory resilience describes the institutional measures (e.g., rules, regulations, and laws) designed to address ecological pressures. The regulatory resilience datasets are grouped into five categories that address the 5 pressure categories: fishing pressure, habitat destruction, climate change, water pollution, and alien species. For this assessment we use two resilience layers for each pressure category where possible: regulations and management.  

Regulation layers are based on if there are established rules to prevent ecological pressures. For alien species, we use the presence of rules to prevent the introduction and spread of invasive species from boats, vessels, and cargo. Fishing and habitat regulations are based on the establishment of protected areas. For fishing resilience, we use the presence of marine protected areas and for habitats, we consider both terrestrial protected areas and marine protected areas. Due to the lack of permanent residents, pollution and climate change don't have localized rules and therefore aren't considered as resilience layers.

Management layers are derived for climate change, habitats, fishing, and alien species based on evaluation of three questions:   
 1. Is there a management plan?    
 2. Does there exist some sort of monitoring or tracking of the effects of the management plan?   
 3. Has the management plan been effective in changing behaviors or meeting its objectives?   
 
A score of one is given for every "Yes" response, a zero given for a "No" response, and 0.5 given if management plans or monitoring programs are "In Development". The overall resilience layer score is the sum over three. Since the inverse of the habitat management resilience is used for a social pressure, we will consider this a social resilience instead of an ecological one.    

The French Polynesian government has a variety of regulations for incoming vessels of all sizes. They are outlined in the 
Stopover Handbook for French Polyesia for Mariners, 2020-2021. From this we know that there are strong protections for many categories of species year round, as well as detailed directions for incoming mariners on registration, mooring, waste management, etc. 

Tetiaroa is not specifically covered in the handbook, due to the fact that it is privately owned. The map on Page 2 does indicate that there is restricted fishing in the area, but we do not know exactly what these are. Therefore, for now we will create this resilience layer based on the regulations that apply nation wide.

## Fishing Regulations  

The following classess of species are prohibited from being captured, possesed, or eaten year round:
- Sharks (referred to [this article](https://www.sciencedirect.com/science/article/pii/S0308597X16305899) for more details on shark restrictions in FP)
- Marine mammals
- Rays
- Troca, burgau, conch, giant mussel, and triton shells
- Black coral 

Some species have seasonal restrictions (lobsters) and size requirements, but we don't have a full list of those. 


# Set Up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)

source(here('workflow/R/common.R'))


dir_resilience <- file.path(dir_github,"prep/resilience")
```

# Methods

For the species regulations, we will calculate the number of species that have strict, year-round regulations as a percentage of all species found in the IUCN global range pull. 

## Read in species status layer 

```{r}

species <- read_csv(here("/prep/bd/spp/v2020/output/spp_status.csv"))

```

## Find species with protections

This layer already has each species class listed, so we can use that to pull out species that fall into any of the protected categories outlined above. For shells and coral, we'll have to look for that specific species. 

Note: the yellow-bellied sea snake is included  in reptilia, need to filter it out

```{r}

protected_class <- c("chondrichthyes", "mammalia", "reptilia")

protected_sp <- species %>% 
  filter(class %in% protected_class) %>% 
  filter(common != "yellow-bellied sea snake")

```


## Option 1: Find percentage of total 

We see that there are 59 species that are heavily restricted, out of 657 species total. If we think this is the best way to measure resilience, then we can just calculate a simple percentage:

```{r}

res_fishing_reg_1 <- data.frame(
  region_id = 1, 
  year = seq(2014,2020,1),
  res_score = 59/657
)

#using the other version

res_fishing_reg_2 <- data.frame(
  region_id = 1, 
  year = seq(2014,2020,1),
  res_score = (1 + 0 + 1)/3
)

#write to folder

write_csv(res_fishing_reg_1, file.path(dir_resilience, "ecological/fishing_reg/res_fishing_reg.csv"))

#write to toolbox

write_csv(res_fishing_reg_2, file.path(dir_scores, "region/layers/res_fishing_reg.csv"))
```

