---
title: "global coral distributions"
author: "Gage Clawson"
date: "7/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

This script generates the extent of coral for each Tetiaroa.  

***
## Data Source 

**Downloaded**: 07/25/2019

**Description**:  
Global Distribution of Coral Reefs
https://data.unep-wcmc.org/datasets/1
Reported at spatial cell scale. 

This dataset shows the global distribution of coral reefs in tropical and subtropical regions. It is the most comprehensive global dataset of warm-water coral reefs to date, acting as a foundation baseline map for future, more detailed, work. This dataset was compiled from a number of sources by UNEP World Conservation Monitoring Centre (UNEP-WCMC) and the WorldFish Centre, in collaboration with WRI (World Resources Institute) and TNC (The Nature Conservancy). Data sources include the Millennium Coral Reef Mapping Project (IMaRS-USF and IRD 2005, IMaRS-USF 2005) and the World Atlas of Coral Reefs (Spalding et al. 2001).

**Time range**: 1954-2018


***
# Methods
Reclassify the coral extent data into a mask of 1 or NA, and then compute zonal statistics for the count of cells within Tetiaroa that have coral and then convert into km2.


## Setup
``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(readr)      # for read_csv()
library(raster)
library(here)
library(sf)
library(fasterize)
library(tidyverse)

source(file.path('~/github/ohiprep_v2020/workflow/R/common.R'))

goal     <- 'globalprep/hab_coral/v2019'
dir_git  <- file.path('~/github/ohiprep_v2019', goal)
dir_wcmc <- file.path(file.path(dir_M, 'git-annex/globalprep/_raw_data/wcmc_coral'))
ohi_rasters() # call the region zones raster
```

```{r, echo = FALSE, eval = FALSE}

v4_coral_pts <- sf::st_read(dsn = file.path(dir_wcmc, "14_001_WCMC008_CoralReefs2018_v4/01_Data"), layer = "WCMC008_CoralReef2018_Pt_v4")

v4_coral_py <- sf::st_read(dsn = file.path(dir_wcmc, "14_001_WCMC008_CoralReefs2018_v4/01_Data"), layer = "WCMC008_CoralReef2018_Py_v4")

v4_coral_metadata <- sf::st_read(dsn = file.path(dir_wcmc, "14_001_WCMC008_CoralReefs2018_v4/01_Data"), layer = "Metadata_CoralReefs")


```

**Convert seagrass shapefiles into same CRS as our region zones raster**
```{r, echo = FALSE, eval = FALSE}

moll_crs <- crs(zones, asText = TRUE)

v4_coral_pts_moll <- st_transform(v4_coral_pts, crs = moll_crs) #project points shapefile to match zones crs


crs(v4_coral_pts_moll) #check to see it worked

v4_coral_py_moll <- st_transform(v4_coral_py, crs = moll_crs)


## Tetiaroa extent
azim_eq    <- raster::crs("+proj=laea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")
azim_ext <- raster::extent(-6156061, -6136000,-4607084,-4588580) #this is for equal area projection only
plot(azim_ext)

global_py <- v4_coral_py_moll %>%
  st_transform(crs = azim_eq)  

tet_py <- st_crop(global_py, azim_ext)

summary(tet)
plot(tet[21])
colnames(tet)

tet_py_1 <- st_cast(tet_py$geometry, "MULTIPOLYGON") 

st_geometry(tet_py) <- tet_py_1 # fix all geometries to be consistent

plot(tet_py[21])

global_pt_eq <- v4_coral_pts_moll %>%
  st_transform(crs = azim_eq)
tet_pt <- st_crop(global_pt_eq, azim_ext) ## nothing in the points data. Will only work with polygons then...


tet_rast <- raster::raster("../../../../spatial/rasters/ocean_ras_100.tif")
plot(tet_rast)
```

**Fasterize/rasterize: Where there is coral assign a value of 1 and NA otherwise**
```{r}
#### fasterize the polygons to our 3nm raster
fasterize_tet_py <- fasterize::fasterize(tet_py, raster = tet_rast, 
                                     field = NULL)
plot(fasterize_tet_py)
#### check to see if all points are NA and 1 for polygon raster
check_values_py <- getValues(fasterize_tet_py)
sum(check_values_py == 1, na.rm = TRUE)
# 2565
sum(is.na(check_values_py))
# 34620
unique(check_values_py)
# NA 1
```

**Calculate zonal stats with 3nm raster and fasterized Palmyra polygons. Convert to km^2**
```{r, echo = FALSE, eval = FALSE}
zonal_sums <- raster::zonal(fasterize_tet_py, 
                                     tet_rast,
                                     fun = "sum",
                                     na.rm = TRUE) #sum all coral cells for each ohi zone

zonal_sums_df <- data.frame(zonal_sums)

my_raster_res <- res(fasterize_tet_py)

my_raster_res ## 100 100
# 100 m to km = 0.1 km sized cells

zonal_sums_km2 <- zonal_sums_df %>%
  mutate(year = 2002, habitat = "coral",
         km2 = (0.1**2*sum)) %>% #one cell is equal to ~0.1 km
  dplyr::rename("rgn_id" = "zone") %>%
  dplyr::select(-sum)

## this shows 24.6 km2 compared to calculated GIS area in the raw data of 2294.982 km2...

sum(tet_py$GIS_AREA_K)
```
