---
title: "get_coral_data"
output: html_document
---

#Summary

This script pulls more up to date coral data from the Living Oceans Foundation French Polynesia report, which was published in 2017. This will be used to assess current status of benthic habitats on Tetiaroa

***

#Data source

**Downloaded**: 07/10/2020

**Description**:  
French Polynesia Final Report
https://data.unep-wcmc.org/datasets/

Description

**Time range**: 2017

#Set up 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(pdftables)
library(here)

source(here('workflow/R/common.R'))

#file paths
goal     <- 'hab'
scenario <- 'v2020'
#dir_anx       <- file.path(dir_M, 'git-annex/globalprep') 
dir_goal      <- file.path(here('prep/bd',goal, scenario))
#dir_goal_anx  <- file.path(dir_anx, goal, scenario)
dir_github    <- '~/github/tet-prep'

pdf_api <- "yclryli65jdy"
```

Use pdftables to convert the data on page 21 of the pdf
```{r}

convert_pdf(file.path(dir_goal,"_raw_data/French-Polynesia-pg-21.pdf"), "fp_habtable.csv", api_key = pdf_api)

```

This worked - had to export and clean it up, resaved as fp_habtable_clean

```{r}
hab_table <- read_csv(here("prep/bd/hab/v2020", "fp_habtable_clean.csv")) %>% 
   mutate(
    area_km2 = as.numeric(Tetiaroa)
  ) %>% 
  dplyr::select(habitat, classification,  area_km2, -Tetiaroa)
 

```

### Habitat sums 

We can now find sums of each type of habitat. The classification also allows us to find coral / algae ratios.
```{r}

hab_stats <- hab_table %>%
  filter(!is.na(area_km2)) %>% 
  group_by(classification) %>%
  summarize(
    tot_area_km2 = sum(area_km2)
  )
#export this as a csv, may be helpful to send to someone

write_csv(hab_stats, file.path(dir_goal, "int","hab_total_area.csv"))

```

### Find the total area of coral cover in km2

We'll treat the categories that have the word "coral" or "reef" as area with coral cover
```{r}

coral_stats <- hab_stats %>% 
  filter(str_detect(classification, pattern = "coral") | str_detect(classification, pattern = "reef")) %>% 
  summarize(
    total_area_km2 = sum(tot_area_km2)
  ) %>% 
  mutate(
    class = "coral extent",
    year = 2017
  )


```

