---
title: "OHI Tetiaroa: Atoll Habitat and Vegetation"
output:
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3 
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/tet-prep/workflow/templates/tet_hdr.html'
  pdf_document:
    toc: true
---

# Summary

This script calculates the land habitat extent, health and trned. Because a long term goal for Tetiaroa is to restore abandoned coconut plantation areas to native vegetation, the health score of the land habitat will be based on the extent of coconut grove versus native habitat. We will also include altered habitat (ie landscaped) since much of the habitat on Onetahi has been altered for construction of the Brando. 

Possible data used this goal comes from a variety of sources, including a 2018 forest monitoring project undertaken by researchers from the University of French Polynesia and Oxford University (McIntosh et al., 2018) in addition to the CASUP_vegetation_flora_JYM_2015.docx found in the dropbox. However, Frank felt that neither of these documents provided an accurate enough assessment of vegetation on each motu. The monitoring report only includes data for small plots on each motu, and the document just lists which habitats exist on each motu, not their extents.

Frank used Google Earth to estimate the area of coconut grove, coconut grove and native habitat mix, and just native habitat. He also includes the estimated area of landscaped habitat on Onetahi. 

We will assign scores based on whether the motu is dominated by native vegetation or by coconut grove. Each motu will receive a score, and then the final score will be calculated using the average weighted by area of the motu. 


# Set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(sf)
library(here)
library(janitor)

source(here('workflow/R/common.R'))

#file paths
goal     <- 'hab'
scenario <- 'v2020'
dir_goal      <- file.path(here('prep/bd',goal, scenario))
dir_goal_anx  <- file.path(dir_anx, goal, scenario)
dir_github    <- '~/github/tet-prep'

#read in habitat data

tet_hab <- read_csv(file.path(dir_goal, "_raw_data/frank_motu_veg_type.csv")) %>%
  clean_names()
```

# Methods

## Compare areas

First, we'll just makes sure that the areas calculated with our shapefile are similar to the areas Frank has on his document.

```{r}

frank_area <- tet_hab %>%
  dplyr::select(name = motu,total_area_ha = total_area)

area_comparison <- left_join(tet_motus, frank_area, by = "name") %>%
  filter(name != "Aie") %>% 
  mutate(
    area_calc = total_area_ha * 0.01
  )
  
## looks pretty good


```

## Extent Native Habitat 

Sum up the extent of native habitat on Tetiaroa. We will include the coco / native habitat and multiply it by 0.5

```{r}

native_area <- hab_scores <- tet_hab %>%
  dplyr::select(motu,coco_area,native_coco_area,native_area,landscape_area,total_area) %>%
  pivot_longer(cols = c("coco_area","native_coco_area","native_area","landscape_area","total_area"), names_to = "habitat",values_to = "area_ha") %>% 
  filter(habitat == "native_area" | habitat == "native_coco_area") %>% 
  group_by(habitat) %>% 
  summarize(
    total_area = sum(area_ha)
  ) %>% 
  mutate(
    total_area = case_when(
      habitat %in% "native_coco_area" ~ 0.5*total_area,
      TRUE ~ total_area
    )
  ) %>% 
  summarize(
    total_native_km2 = sum(total_area)*0.01 #convert to km2
  )

native_extent_df <- data.frame(
  region_id = 1,
  habitat = "native_land",
  year = 2020,
  km2 = native_area$total_native_km2
)

#write to tool box

write.csv(native_extent_df, file = "~/github/tet-scores/region/layers/hab_land_extent.csv", row.names = FALSE)

```


## Assigning habitat scores to determine health of habitat

Looking at the CASUP habitat document, we can see that there are four motus that have "very dense coco forest". All four also have another type of vegetation. Do some exploring to come up with the best scoring scheme: 

We'll use what Frank has for this calculation, but information from the forest study helps underscore which native trees should be present. The summary below includes information about these species. Note: if the tree is designated on the Tetiaroa website as 'Polynesian Introduction' we'll still consider it native.

Below are the results from the study: 

| Motu | Dominant species | Native? | Notes |
| :----- | :--------- | :----- | :-------------------|
| Reiono | Pisonia grandis / Cabbage tree | yes | Important nesting site, helps protect against storm surge |
| Tiaraunu N | Cocos nucifera / Coconut palm | no | Coco plantations have made motus more vulnerable to storm surge and coastal erosion
| Tiaraunu S | Callophyllum inophyllum / Tamanu | yes | Introduced by native Polynesians, oil from its fruit used for many purposes |
| Rimatu | Pandanus tectorius / Screwpine | yes | Key food staple for native atoll inhabitants, used also for medicinal purposes and weaving | 
| Onetahi | Casuarina equisetifoilia / Ironwood | yes | Introduced by native Polynesians, can grow to 100 ft tall, symbol of the war god 'Oro |
| A'ie | Mixed forest | yes | No further details |


Between the report and the species profiles on the Tetiaroa Society we can see that the native ecosystem has many benefits compared to coconut groves. Also, while landscaped areas are most likely replacing removed or altered native habitat, we know from The Brando's website that they only use native plants and do their best to recreate the ecosystem. We therefore will score these areas slightly better than "only coconut grove". See full scoring scheme below:

| Habitat | Score |
| :--------------- | :------------ |
| Only native vegetation | 1 |
| Native vegetation mixed with coconut grove | 0.5 |
| Landscaped area | 0.3 |
| Only coconut grove | 0 | 


## Calulate the habitat health
```{r}

native_health <- tet_hab %>%
  dplyr::select(motu,coco_area,native_coco_area,native_area,landscape_area,total_area) %>%
  pivot_longer(cols = c("coco_area","native_coco_area","native_area","landscape_area","total_area"), names_to = "habitat",values_to = "area_ha") %>% 
  group_by(habitat) %>% 
  summarize(
    total_area_ha = sum(area_ha)
  ) %>%
  ungroup() %>%
  pivot_wider(names_from = "habitat", values_from = "total_area_ha") %>% 
  mutate(
    health = (native_area + 0.3*landscape_area + 0.5*native_coco_area)/total_area
  )

native_health_df <- data.frame(
  region_id = 1,
  habitat = "native_land",
  year = 2020,
  health = native_health$health
)

#write to tool box

write.csv(native_health_df, file = "~/github/tet-scores/region/layers/hab_land_health.csv", row.names = FALSE)

```

## Calculate native habitat trend

Since this makeup has not changed drastically in the last 5 years, the trend will be zero

```{r}

#create trend df for toolbox

years = seq(2014,2020,1)

native_trend <- data.frame(
    region_id = rep(1, length(years)),
    habitat = rep("native_land", length(years)),
    year = years,
    trend = rep(0, length(years))
)

native_trend_sm <- data.frame(
  region_id = 1,
  habitat = "native_land",
  year = 2020,
  trend = 0
)

write.csv(native_trend_sm, file = "~/github/tet-scores/region/layers/hab_land_trend.csv", row.names = FALSE)
```
