---
title: 'OHI Tetiaroa: Fishing Regulations Resilience Layer'
author: 'Madeline Berger'
output:
  html_document:
    highlight: haddock
    includes:
      in_header: '~/github/tet-prep/workflow/templates/ohi_hdr.html' 
    number_sections: yes
    theme: cerulean
    toc: yes
---
# Summary 

The French Polynesian government has a variety of regulations for incoming vessels of all sizes. They are outlined in the 
Stopover Handbook for French Polyesia for Mariners, 2020-2021. From this we know that there are strong protections for many categories of species year round, as well as detailed directions for incoming mariners on registration, mooring, waste management, etc. 

Tetiaroa is not specifically covered in the handbook, due to the fact that it is privately owned. The map on Page 2 does indicate that there is restricted fishing in the area, but we do not know exactly what these are. Therefore, for now we will create this resilience layer based on the regulations that apply nation wide.

## Fishing Regulations  

The following classess of species are prohibited from being captured, possesed, or eaten year round:
- Sharks (referred to [this article](https://www.sciencedirect.com/science/article/pii/S0308597X16305899) for more details on shark restrictions in FP)
- Marine mammals
- Rays
- Troca, burgau, conch, giant mussel, and triton shells
- Black coral 

Some species have seasonal restrictions (lobsters) and size requirements, but we don't have a full list of those. 

## Biosecurity 

All incoming vessels must declare their plants, plant products, and animals before unboarding. Certain products are prohibited, like honey. Vessels from certain countries with known threats (ie coconut rhinocerous beetle) all must be inspected. 

Offenders of these rules face up to 6 months in prison and a fine of up to 3.57 million Francs. 


# Set Up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)

source(here('workflow/R/common.R'))


dir_resilience <- file.path(dir_github,"prep/resilience")
```

# Methods

For the species regulations, we will calculate the number of species that have strict, year-round regulations as a percentage of all species found in the IUCN global range pull. 

## Read in species status layer 

```{r}

species <- read_csv(here("/prep/bd/spp/v2020/output/spp_status.csv"))

```

## Find species with protections

This layer already has each species class listed, so we can use that to pull out species that fall into any of the protected categories outlined above. For shells and coral, we'll have to look for that specific species. 

Note: the yellow-bellied sea snake is included  in reptilia, need to filter it out

```{r}

protected_class <- c("chondrichthyes", "mammalia", "reptilia")

protected_sp <- species %>% 
  filter(class %in% protected_class) %>% 
  filter(common != "yellow-bellied sea snake")

```


## Option 1: Find percentage of total 

We see that there are 59 species that are heavily restricted, out of 657 species total. If we think this is the best way to measure resilience, then we can just calculate a simple percentage:

```{r}

res_fishing_reg_1 <- data.frame(
  region_id = 1, 
  year = seq(2014,2020,1),
  res_score = 59/657
)

#write to folder

write_csv(res_fishing_reg_1, file.path(dir_resilience, "ecological/fishing_reg/res_fishing_reg.csv"))
```

