---
title: "oa_prep_mb"
author: "Maddie Berger"
date: "8/10/2020"
output: html_document
---
# Summary 

This ocean acidification (OA) pressure layer preparation script does the following for newly available OA data:

- Reads in the annual means rasters from OHI global
- Resamples the rescaled raster layer to (1km^2^ cell resolution) the same resolution as the Tetiaroa ocean raster
- Masks the resampled data to select only those cells within the ocean

***

# Data Source

**Reference**: [Feely et al.(2009)](https://darchive.mblwhoilibrary.org/bitstream/handle/1912/3180/22-4_feely.pdf?sequence=1&isAllowed=y)

**Downloaded**: July 19, 2017

**Description**: Aragonite Saturation State  $\Omega_{arg}$

**Native data resolution**: 1 degree cells

**Time range**: 1880-1889 and 1958-2100, monthly data provided for each year. Future years are based on model projections for RCP 8.5. Previous years are hindcast/historical data.

**Format**: NetCDF

**Citation information**: Woods Hole Oceanographic Institution. 2014 update to data originally published in: Feely, R.A., S.C. Doney, andvS.R. Cooley. 2009. Ocean acidification: Present conditions and future changes in a high-CO2 world. Oceanography 22(4):36–47

**Notes about the data**: This data was shared with NCEAS by Ivan Lima from Woods Hole Institute for Oceanography in July 2017. Originally, data were provided from 1958 - 2016; additional years of data up until 2020 were received from Ivan Lima. The data came as NetCDFs with a resolution of about 1 degree. The data values are monthly average surface &#937; aragonite saturation state.

***

# Setup 

```{r setup, message = F, warning = F}

# Set options for all chunks in code
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.width=6, fig.height=6)

library(raster)
library(ncdf4)
library(maps)
library(parallel)
library(foreach)
library(doParallel)
library(RColorBrewer)
library(fasterize)
library(rasterVis)
library(here)
library(gstat)
library(scales)
library(animation)

# Source and set file paths, will also load Tetiaroa shapefiles and rasters
source(here('workflow/R/common.R'))

raw_dir <- file.path(dir_M, 'git-annex/globalprep/_raw_data')

dir_github <- here()
```

# Method 1 - Zonal Stats from final Global Output

## Read in global data from v2017 output 

The data from this source are: 
- resolution = 1 km2
- its been rescaled 0 - 1 based on historical and biological reference points (see XXXX for more details) 

```{r}

rasts <- c(list.files(file.path(dir_M, "git-annex/globalprep/prs_oa/v2017/output"),full.names=T))

three_nm <- raster(file.path(dir_github, "spatial/rasters/ocean_ras_100.tif"))
three_nm <- raster::crop(three_nm, azim_ext) # Crop to Tetiaroa extent
plot(three_nm)

# if you need more processsing power:

registerDoParallel(8)

# this loop resamples the global output layers to teitaroa extent and resoltuion. 

foreach(file = rasts) %dopar% {

  #file = rasts[3]
  yr = substr(file, 72,75)

  rast_azim <- raster(file) %>%
    projectRaster(from = rast_azim, to = three_nm, crs = "+proj=laea +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs", method = "ngb") %>%
     writeRaster(., filename = paste0(file.path(dir_anx), '/prs_oa/int/global_rescaled_azimuthal/global_rescaled_cropped_', yr, '.tif'), overwrite=T)
}

# grab these new files and put into a stack
rasts_crop <- c(list.files(file.path(dir_anx, "prs_oa/int/global_rescaled_azimuthal"),full.names=T))

plot(raster(rasts_crop[5]), col=cols,box=F,axes=F)
pressure_stack <- stack(rasts_crop)

```

## Get the mean value for the Tetiaroa zone

```{r}

regions_stats <- zonal(pressure_stack,  tet_ocean_rast_100, fun="mean", na.rm=TRUE, progress="text")
write.csv(regions_stats, file.path(dir_github,"/prep/pressures/oa/v2020/int/region_stats.csv"), row.names=FALSE)

#clean up the data frame
region_stats2 <- data.frame(regions_stats) %>% 
  pivot_longer(cols = starts_with("global"), names_to = "year", values_to = "pressure_score") %>% 
  filter(year != "global_rescaled_cropped_ed_c")

oa_data <- region_stats2 %>% 
  mutate(
    year = as.numeric(substr(year,25,28))
  )

#write final to output

write.csv(oa_data, file.path(dir_github,"/prep/pressures/oa/v2020/output/oa_method_one.csv"), row.names=FALSE)

```

**Note: these all came out the same for every year also**


# Method 2 - Zonal stats from Raw Data

This method starts with the raw rasters from global, reprojecting and calculating zonal stats all in one loop. We then need to rescale using the reference points. 

## Read in the raw data
```{r}

# Read in 1958-2016 data
raw <- stack(file.path(raw_dir, 'WHOI/cesm_co2sys_1958-2016.1x1d.nc'), varname="OARG") # Grab aragonite variable using varname
                    
plot(raw[[1]]) # We'll have to rotate the rasters
plot(rotate(raw[[1]]))
maps::map('world', col='gray95', fill=T, border='gray80', add=T)


```

## Reproject, crop in this loop:

```{r}

# Register 10 cores for parallel processing
registerDoParallel(10)

foreach (i = seq(1,708, by=12)) %dopar% { 
  #i = 1

  j = i+11
  
  yr = substr(names(raw[[i]]),2,5)
  
  yr_mean = raw[[i:j]] %>%
            calc(fun=function(x){mean(x, na.rm=T)}) %>%
            rotate() %>%
            projectRaster(tet_ocean_rast_100, method = "ngb") %>% # Reproject to Tetiaroa ocean raster
            crop(azim_ext) %>%
            #resample(hist, method = 'ngb') %>% 
             #returning a blank raster
          writeRaster(., filename = paste0(file.path(dir_anx), '/prs_oa/int/annual_avg_azimuthal/tet_arag_avg_azimuthal_', yr, '.tif'), overwrite=T)
 
}

#read in these files

rasts <- c(list.files(file.path(dir_anx, "prs_oa/int/annual_avg_azimuthal"),full.names=T))

stack <- stack(rasts)

```

Find zonal statistics 

```{r}

zonal <- zonal(stack, tet_ocean_rast_100, fun = "mean", na.rm= T)
write.csv(zonal, file.path(dir_github,"/prep/pressures/oa/v2020/int/region_stats_2.csv"), row.names=FALSE)

region_stats3 <- data.frame(zonal) %>% 
  pivot_longer(cols = starts_with("tet"), names_to = "year", values_to = "pressure_score") %>% 
  mutate(
    year = substr(year, 24, 27)
  )

write.csv(region_stats3, file.path(dir_github,"/prep/pressures/oa/v2020/int/oa_method_two_raw.csv"), row.names=FALSE)
```

Could we just repeat this with the historical reference point? or can we just grab the historical reference point from global and use a quick mutate to rescale?

## Get historical baseline - crop, reproject and get zonal stats

```{r}

hist <- raster(file.path(dir_M, 'git-annex/globalprep/prs_oa/v2015/working/global_oa_1880_1889_arag_mean_moll.tif'))

plot(hist, main='Mean Ωaragonite 1880-1889', col=rev(cols), box=F, axes=F)


hist_tet <- hist %>%
  projectRaster(tet_ocean_rast_100, method = "ngb") %>% # Reproject to Tetiaroa ocean raster
  crop(azim_ext)
  #resample(hist, method = 'ngb') %>% 
  #writeRaster(., filename = paste0(file.path(dir_anx), '/prs_oa/int/annual_avg_azimuthal/tet_arag_avg_azimuthal_', yr, '.tif'), overwrite=T)
  

plot(hist_tet)

zonal_hist <- zonal(hist_tet, tet_ocean_rast_100, fun = "mean", na.rm= T)
region_hist <- data.frame(zonal_hist)
```

The historical mean is 4.67. Apply this to the region_stats3 df to get the difference and scale it. Use the following equation: 

>This pressure layer is rescaled so that all values lie between 0 and 1 using both a historical reference period and >a biological reference point. All cells with values less than 1, indicating an undersaturated state, are set equal >to the highest stressor level, 1. For all other cells, rescaling the aragonite saturation state value to between 0 >and 1 relies upon the change in saturation relative to the reference period.

>Deviation from aragonite saturation state is determined for each year in the study period using this equation:

$$\Delta \Omega_{year} = \frac{(\Omega_{base} - \Omega_{year})}{(\Omega_{base} - 1)}$$

```{r}

tet_oa <- read_csv(file.path(dir_github,"/prep/pressures/oa/v2020/int/oa_method_two_raw.csv"))

hist <- 4.67

tet_oa_rescaled <- tet_oa %>%
  mutate(
    mean = pressure_score,
    diff = (hist-mean)/(hist-1),
    diff2 = hist-mean
  ) 
  
  
#check this - all less than one so are these just one then?
```




