---
title: "Tetiaroa RRedlist"
author: "Maddie Berger"
date: "6/4/2020"
output: html_document
---

#Summary

Preparing data layer for iconic species, based on list from the Tetiaroa Society website. Thanks to Gage Clawson for helping out with the RRedlist package!

#Set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rredlist)
library(tidyverse)
library(dplyr)
library(raster)
library(here)
library(janitor)
source(here('workflow/R/common.R'))

##### api keys for IUCN #####

api_file <- file.path(dir_M, 'git-annex/globalprep/spp', 
                      'api_key_gc.csv')
api_key <- scan(api_file, what = 'character')

'%!in%' <- function(x,y)!('%in%'(x,y)) #define helpful filtering function


##### file paths #######

goal     <- 'ico'
scenario <- 'v2020'
dir_anx       <- file.path(dir_M, 'git-annex/globalprep') 
dir_goal      <- file.path(here('prep/sp',goal, scenario))
dir_goal_anx  <- file.path(dir_anx, goal, scenario)
dir_github    <- '~/github/tet-prep'


##### read in raw data ######

tet_species <- read_csv(here("./prep/sp/ico/v2020/raw_data", "tet_species.csv"))%>%
  clean_names() %>% 
  mutate(
    scientific_name = str_to_lower(scientific_name) #for easier matching, will also do on the redlist df
  )
```

#Method

##Part 1: Filtering Redlist by Location (French Polynesia)
Get US Minor French Polynesia species 
```{r}
spp_country_url <- 'http://apiv3.iucnredlist.org/api/v3/country/getspecies/id/%s?token=%s'
# 
# /api/v3/country/getspecies/:country?token='YOUR TOKEN'
countries <- rl_countries(key = api_key)$results
#UM is United States Minor Outlying Islands country code
# UM species risks
spp_um <- rl_sp_country('UM', key = api_key)$result %>%
  mutate(country_code = "UM") %>%
  rename("iucn_sid" = "taxonid")
##PF is French Polynesia country code
## French Polynesia Species risks
spp_fp <- rl_sp_country('PF', key = api_key)$result %>%
  mutate(country_code = "PF") %>%
  rename("iucn_sid" = "taxonid") %>% 
  mutate(
    scientific_name = str_to_lower(scientific_name)
  )

```

Filter out species we want for Tetiaroa:
```{r}

tet_species_iucn <- spp_fp %>% 
  filter(scientific_name %in% tet_species$scientific_name)

tet_species_iucn_fp <- inner_join(tet_species_iucn,tet_species, by = "scientific_name") %>%  dplyr::select(iucn_sid, scientific_name,category.x,name,ico)

```

Check to see if they were all there:
```{r}

missing_spp <- setdiff(tet_species$scientific_name, tet_species_iucn$scientific_name)

tet_species_missing_fp <- tet_species %>% 
  filter(scientific_name %in% missing_spp)
```

Seems like the French Polynesia list is missing some species that are found on Tetiaroa
https://rdrr.io/cran/rredlist/man/rl_occ_country.html

```{r}

#test hawksbill turtle, which is critically endangered

rl_occ_country("Eretmochelys imbricata", key = api_key) 
#can see that FP isn't on here, even though the range maps shows it occurs there

```

##Part 2: Using the entire IUCN dataset, instead of filtering by country

Using the IUCN API, we accessed the full IUCN species list at http://apiv3.iucnredlist.org/api/v3/speciescount?token=. With some minor formatting, this list contains the following variables:

iucn_sid | kingdom | phylum | class | order | family | genus | sciname | population | category


```{r}
## get a dataframe of all iucn redlist species

out <- rl_sp(all = TRUE, key = api_key)

all_df <- do.call(rbind, lapply(out, "[[", "result")) #creates data frames from each item in the large list in `out` and then binds them

all_df_comp <- all_df %>%
  dplyr::select(-infra_rank, -infra_name) %>%
    dplyr::rename(iucn_sid = taxonid, sciname = scientific_name) %>% 
    setNames(names(.) %>%
               stringr::str_replace('_name', '')) %>% 
  mutate(
    sciname = str_to_lower(sciname)
  )


#filter large df by the all the scientific names vector

tet_listed <- all_df_comp %>% 
filter(sciname %in% tet_species$scientific_name)

#this only has 86 obs, improvement from before

not_in_iunc <- setdiff(tet_species$scientific_name, tet_listed$sciname) 

no_record <- tet_species %>% 
  dplyr::filter(scientific_name %in% not_in_iunc) #49 observations

#this seems really unlikely. something might be up with spelling?


```

So far, have created the following dfs:
- `tet_species_iucn_fp` is all the species that matched what was pulled from the IUCN French Polynesia list
- `tet_species_missing_fp` is all the species that were missing from the IUCN French Polynesia list but were included in Tetiaroa's species list
- `tet_listed` is all the species that were matched using the entire IUCN dataset, which gets us up to 86 - a bunch of duplicates though, need to filter out by range
- `no_record` includes all the species we still can't match. This list is probably a result of differences in spelling as there are many species on there that should be in the IUCN, like humpback whale, spotted eagle ray, and olive ridley's turtle. 

###Working with `tet_listed` dataset

This dataset needs to be cleaned up and filtered out for the sub-populations we are interested in, since the IUCN category differs for them. 

```{r}

tet_lc <- tet_listed %>% 
  dplyr::select(iucn_sid, sciname, population, category)

#find the duplicate species

tet_listed_dup <- tet_lc[duplicated(tet_lc$sciname),]

#try by just filtering out the ones with key words in the population columns - these are the versions we want!

test <- tet_lc[grepl("South Pacific", tet_lc$population) | grepl("Hawaiian", tet_lc$population) | grepl("Asia", tet_lc$population)| grepl("East Pacific", tet_lc$population) ,]

#now filter out the duplicated species from the tet_lc dataset, and bind the test data set to it (probably a better way of doing this)

tet_lc_2 <- tet_lc %>% 
  filter(sciname %!in% tet_listed_dup$sciname)

tet_listed_clean <- bind_rows(tet_lc_2, test)

#61 matched species

tet_listed_info <- tet_species %>% 
  dplyr::select(name, sciname = scientific_name)

tet_listed_ico <- inner_join(tet_listed_info, tet_listed_clean, by = "sciname")

```

###Working with `no_record` dataset

Now we can work on the other batch, the ones that did NOT match. While 61 iconic species is plenty, there are a few key species we want to add that seem like obvious oversights. These include:  

- **Humpback whale**: iucn spells it "megaptera novaeangliae", which is the same? IUCN id = 132832
- **Melon headed whale**: misspelled IUCN id = 16564
- **Olive Ridley sea turtle**: misspelled IUCN id = 11534
- **Clown fish**: correct spelling, species specific to Tetiaroa not included in IUCN data set
- **Spotted eagle ray**: spelled "aetobatus narinari", IUCN ID = 39415
- **Pink whipray**: not found in IUCN data set
- **Strawberry hermit crab**: spelled correctly, not included
- **Christmas tree worm**: not included

```{r}

#create vector of the ids we want
iucn_sid <- c(132832, 16564, 11534, 39415) 
names <- c("humpback whale", "melon-headed whale", "olive ridley sea turtle", "spotted eagle ray")

miss_info <- data.frame(iucn_sid,names)

#filter entire IUCN dataset, match columns to tet_listed_clean
missing_ico <- all_df_comp %>% 
  filter(iucn_sid %in% iucn_sid) %>% 
  dplyr::select(iucn_sid, sciname, population, category)

#join to common names

missing_ico_clean <- inner_join(miss_info,missing_ico, by = "iucn_sid") %>% 
  dplyr::select(name = names, sciname, iucn_sid, population, category)
```

Last step is to bind missing iconic species to the listed ones: 

```{r}

tet_ico_all <- bind_rows(tet_listed_ico,missing_ico_clean) %>% 
  mutate(
    name = str_to_lower(name)
  )
#write this into int folder

write_csv(tet_ico_all, file.path(dir_goal,'int/tet_ico_all.csv'))
```

## Historic threat

We accessed the IUCN API to determine past IUCN assessments for each of the identified iconic species: http://apiv3.iucnredlist.org/api/v3/species/history/id/?token=.

Each assessment includes a year and an extinction risk, along with additional information on the assessment.

```{r}

tet_ico_list <- read_csv(file.path(dir_goal, 'int/tet_ico_all.csv'))

ico_ids <- unique(tet_ico_list$iucn_sid)

#use rl_history, create loop that runs it for each id in our iconic species list

for(i in seq_along(ico_ids)){

id = ico_ids[i]
out <- rl_history(id = id, key = api_key) #this comes out as a list
df <- as.data.frame(out)

 if(i == 1) {
   tet_ico_history <- df
 } else {
   tet_ico_history <- rbind(tet_ico_history, df)
 }

}

#write out this data frame to int folder
write_csv(tet_ico_history, file.path(dir_goal,'int/tet_ico_history.csv'))
```

#Export final data layers to output

The toolbox wants `rgn_id`, species `sciname`, and extinction risk category for the basic calculations. Since some regions contain multiple subpops (or parent/subpop) we also include `iucn_sid` to differentiate. This information is included for each `year`, filtered back to the year 2000.

- `ico_spp_iucn_status.csv`: rgn_id,sciname,iucn_sid,year,eval_yr,category

Bind scinames, rgn_id vector, and eval_year vector to this data frame
```{r}
#read in data from int folders
ico_spp_iucn <- read_csv(file.path(dir_goal, "int/tet_ico_history.csv")) %>%
  dplyr::select(iucn_sid = name, year = result.year, category = result.code)

tet_ico_all <- read_csv(file.path(dir_goal, "int/tet_ico_all.csv")) %>% 
  dplyr::select(sciname, iucn_sid)

rgn_id <- rep(1, length(ico_spp_iucn$iucn_sid))
eval_year <- rep(2020, length(ico_spp_iucn$iucn_sid))

#bind and join them all up

ico_spp_iucn_status <- data.frame(rgn_id,ico_spp_iucn,eval_year) %>% 
  inner_join(., tet_ico_all, by = "iucn_sid")

#write csv
write_csv(ico_spp_iucn_status, file.path(dir_goal, "output/ico_spp_iucn_list.csv"))


```

##Calculating status scores

So far we've found the conservation status for these species, now calculate score. In global this is done in the tool box but we'll do it in prep here. I've saved both layers in outputs. 

```{r}

#read in list for all years 

tet_ico_list <- read_csv(here("/prep/sp/ico/v2020/output", "ico_spp_iucn_list.csv"))

#create data frame with scores and add column with inverse of risk score (this is status score)

pop_cat <- data.frame(category = c("LC", "NT", "VU", "EN", "CR", "EX", "T", "CD", "NE", "DD"), 
                      risk_score = c(   0,  0.2,  0.4,  0.6,  0.8,  1.0, 0.6,  0.3,   NA,  NA),
                      stringsAsFactors = FALSE) %>% 
  mutate(
    stat_score = 1 - risk_score
  )

#join the scores data to the list data

tet_ico_stat_sco <- left_join(tet_ico_list, pop_cat, by = "category") %>%
  mutate(stat_score = case_when(
    category %in% c("LR/lc") ~ 1.0,
    category %in% c("LR/nt") ~ 0.8,
    category %in% c("LR/cd") ~ 0.9,
    TRUE ~ stat_score)) %>% 
  dplyr::select(-risk_score)

```

##Backfilling missing years
```{r}

#this gap fills based on the earliest assessment for that species
tet_ico_full <- tet_ico_stat_sco %>%
  dplyr::select(-eval_year) %>%  
  mutate(eval_yr = year) %>% 
  arrange(iucn_sid, year) %>% 
  complete(year = full_seq(year, 1), nesting(iucn_sid)) %>% 
  group_by(iucn_sid) %>% 
  fill(sciname, category, stat_score, eval_yr, rgn_id, .direction = "downup") %>%  ## fills all the way to latest year
  ungroup() %>% 
  filter(year > 2013) %>% 
  dplyr::select(rgn_id, year, sciname, status = stat_score)

```

Adding common names and 2020 - step one is to read in the spp int file, because we already did that there

```{r}

common_names <- read_csv(here("/prep/bd/spp/v2020/int", "2_spp_ico_iucn.csv")) %>% 
  dplyr::select(sciname, name) %>% 
  mutate(
    sciname = str_to_lower(sciname),
    name = str_to_lower(name)
  )

#join to tet_ico_full

status_19 <- left_join(tet_ico_full, common_names, by = "sciname")

#add 2020 data
ico_scores <- status_19 %>% 
  filter(year == 2019) %>% 
  mutate(
    year = 2020
  ) %>% 
  bind_rows(.,ico_scores)

```



May need to add a gapfilling file, which would just indicate which years were gapfilled and which were not. 

Write final scores output

```{r}
#write to outputs folder -full version
write_csv(ico_scores, file.path(dir_goal, "output/ico_scores.csv"), append = FALSE)

#write to tool box
write.csv(ico_scores %>% dplyr::select(region_id = rgn_id, year, name, status), file = "~/github/tet-scores/region/layers/ico_status.csv", row.names = FALSE)
```



