---
title: 'OHI+ Tetiaroa: Resource Access Opportunities'
author: "Madeline Berger"
output:
  html_document:
    highlight: haddock
    includes:
      in_header: '~/github/tet-prep/workflow/templates/tet_hdr.html' 
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Summary 

The resource access opportunities goal captures the access people have to coastal resources, whether or not they actually take fish or harvest from coastal waters. 

Data relevant to Tetiaroa that could inform this goal include things like:
- artisanal fishing catch, number of boats, accessible areas
- recreational fishing
- measures of access or availability: boat rental prices, gas prices

For this assessment, we will use a placeholder layer that can be updated once this type of data becomes more available. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)

source(here('workflow/R/common.R'))

#file paths
goal     <- 'rao'
scenario <- 'v2020'
dir_goal      <- file.path(here('prep',goal, scenario))
dir_goal_anx  <- file.path(dir_anx, goal, scenario)

```


## Create a placeholder layer with NA

```{r}


rao_status_placeholder <- data.frame(
  region_id = 1,
  year = seq(2014,2020,1),
  status = NA
)

#write to toolbox

write_csv(rao_status_placeholder, file.path(dir_scores,"/region/layers/rao_status_placeholder.csv"))

```

