---
title: 'OHI Tetiaroa: Local Employees Resilience Layer'
author: "Maddie Berger"
output:
  html_document:
    highlight: haddock
    includes:
      in_header: '~/github/tet-prep/workflow/templates/tet_hdr.html' 
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
---
# Summary

This script creates a resilience layer based on the number of local employees on Tetiaroa. We are using the number reported in the 2018 Beachcomber sustainability report. 

We think this is a measure of resilience because using local knowledge in resource management often results in positive environmental outcomes. Furthermore, it is important that local people are engaged as stakeholders in decisions about Tetiaroa's land and waters, even if it is owned privately. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)

library(here)
library(tidyverse)

source(here('workflow/R/common.R'))

dir_resilience <- file.path(dir_github,"prep/resilience")
```

# Methods

Page 30 in the report states "80% of staff are Polynesian", which exceeds the governments' requirment for local employment. Therefore, Frank and Neil feel the resilience score should be 100%. 

```{r}

res_local_employees <- data.frame(
  region_id = 1, 
  years = seq(2014,2020,1),
  res_score = 1
)

write_csv(res_local_employees, file.path(dir_resilience, "social/local_employees/res_local_emp.csv"))

#toolbox

write_csv(res_local_employees, file.path(dir_scores, "region/layers/res_local_emp.csv"))

```


