---
title: 'OHI+ Tetiaroa 2020: Tourism Runoff'
author: "Madeline Berger"
output: 
  html_document:
    toc: true
    toc_depth: 2 
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
     in_header: '~/github/tet-prep/workflow/templates/tet_hdr.html' 
  pdf_document:
    toc: true
---

# Summary 

This script aims to calculate the pressures generated from tourism on Tetiaroa. In the future, when more data is available, this could include population data to calculate intertidal trampling, or more details on possible contamination from the hotel. For now, this pressure is expressed using the inverse of the waste management score we calculated in the Tourism Sustainability goal. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)

source(here('workflow/R/common.R'))

goal     <- 'pollution'
scenario <- 'v2020'
dir_goal      <- file.path(here('prep/pressures',goal))
```

# Method 

## Read in tourism layers 

```{r}
#grab just the waste management row
tr_data <- read_csv(here("/prep/tr/v2020/int/tr_category_scores.csv")) %>% 
  filter(category == "waste")

po_tr_pollution <- tr_data %>%
  mutate(
    pressure_score = 1 - score,
    region_id = 1,
    year = 2020
  ) %>% 
  dplyr::select(region_id, year, pressure_score)

```

## Save layer

```{r}

write_csv(po_tr_pollution, file.path(dir_goal, "po_tr.csv"))

write_csv(po_tr_pollution, file.path(dir_scores,"/region/layers/prs_po_tr.csv"))

```

