---
title: "OHI Tetiaroa: Conservation Economy"
output:
  html_document:
    highlight: haddock
    includes:
      in_header: '~/github/tet-prep/workflow/templates/ohi_hdr.html' 
    number_sections: yes
    theme: cerulean
    toc: yes
---
# Summary 

This markdown is the initial data prep for the livelihoods and economy goal. There are very few full time residents on the island, and all of them are employed by The Brando or the Tetiaroa Society. Essentially all of the Tetiaroa Society's activities and expenditures are related to environmental and cultural conservation of the island, so we refer to this goal in this context "Conservation Economy" - (MORE). Also, because a portion of the Tetiaroa Society's income comes from guest fees at the resort, we can make the link between a health ocean that attracts tourism, and generating the funds necessary to help preserve it. 

We use the Tetiaroa's Society fundraising goal as the reference point, which breaks down as follows:

| Item | Yearly Cost | Description | 
| :------- | :--------: | :------------ |
| Operations | $700 K | Covers core staff and facilities maintenance / depreciation | 
| **Mission Programs:** |       |          |
| Conservation & Sustainable Use Plan | $ 250 K 
| Social-Ecological System Observatory | $750 K | Monitoring CASUP, data management, and establishing baselines for science |
| Science Programs | $100 K | Seed grants, equipment, meetings, workshops |
| Cultural and Educational Programs | $100 K |       |
| Blue Climate Initiative | $100 K |       |

**Total Yearly Budget** = $2 M USD. 


To measure the status of this goal, we use the 2018 and 2019 Tetiaroa Society reports for income in 2018 and 2019, and then a measure of $700 K / year for 2020 (per Neil and Frank's estimation). 

# Set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)

source(here('workflow/R/common.R'))


# file paths 
goal     <- 'liv'
scenario <- 'v2020'
dir_anx       <- file.path(dir_M, 'git-annex/globalprep') 
dir_goal      <- file.path(here('prep',goal, scenario))
dir_goal_anx  <- file.path(dir_anx, goal, scenario)
dir_github    <- '~/github/tet-prep'

```

# Method

This calculation is relatively simple, simply summing the income sources in each year to get the total income, and then dividing by the refernce point described above to get the status score for each year. 

```{r}
#read in the data

funding <- read_csv(file.path(dir_goal, "raw_data_","TS_income_18_19_updated.csv")) %>% 
  clean_names()


#add 2020 status from Neil's email

ts_income_20 <- data.frame(
  category = "ts_income",
  year = 2020,
  income = 700000
)

#sum each category to get total income in each year

summary <- funding %>%
  #filter(category != "goal") %>% 
  group_by(category, year) %>%
  summarize(income = sum(cost_usd)) %>% #find sum of each category in each year
  bind_rows(., ts_income_20)
  
#add 


#calculate scores and create status df

CE_status <- summary %>% 
  mutate(
    status = income/2000000
  ) %>% 
  filter(category %!in% c("goal","TS_cost")) %>% 
  mutate(
    region_id = 1
  )
```

Write out layers to output folder and to Toolbox

```{r}

write_csv(CE_status, file.path(dir_goal, "output/CE_status.csv"))


```


With this method, the final status scores are: 


| Year | Yearly Income | Status Score |
| :------ | :-------- | :------- | 
| 2018 | $500641.20 |   | 
| 2019 | $828701.00 |   |
| 2020 | $700000.00 |   |
