---
title: "Tourism sustainability"
author: Madeline Berger
output:
  html_document:
    highlight: haddock
    includes:
      in_header: '~/github/tet-prep/workflow/templates/ohi_hdr.html' 
    number_sections: yes
    theme: cerulean
    toc: yes
---

## Summary 

This script calculates a score for tourism sustainability of [The Brando](https://thebrando.com/) resort, the eco-resort on Onetahi. Data for this goal are descriptive sustainability metrics that I gathered from SA Beachcomber's 2019 sustainability report. Most of them are just in percentages (i.e. 80% energy provided by solar).

We did inquire about actual data from The Brando, contacting Alban Perret (sustainability manager). He was not able to share any data at this time, as The Brando has had to push their sustainability work to 2021. Alban said to reach out again in Q2 2021 for an update. His contact information is: aperret@pbsc.pf

The Brando has already achieved far more in terms of sustainability than most resorts, so setting a reference point is tricky. They were the first resort building to receive LEED Platinum certification, and already have a very low impact. We'll assume that being at 100% sustainable, i.e. no footprint, is a goal. While it does not seem totally feasible, the 2019 sustainability report outlines the following broad goal, which supports this reference point: 

>"Decrease carbon footprint of human activity on the island, working towards zero impact"


There are a handful of green / eco / sustainable tourism certifications out there, but one that provides a helpful framework is the [Green Destinations](https://greendestinations.org/green-destinations-standard/) certification. This certification has 100 criteria that can be grouped into 6 main themes: 

1. Destination Management
2. Nature and Scenery 
3. Environment and Climate
4. Culture and Tradition
5. Social Well-being
6. Business and Hospitality

[EcoTourism Australia](https://www.ecotourism.org.au/our-certification-programs/eco-destination-certification/) considers any place that has met 80% of the requirements in this standard a "nature destination", a place that has met 90% as "ecotourism destination" and one that meets 100% as officially Green Destination Certified. 

For now I'll just use this framework to group the statistics I have into categories. Under Environment and Climate, Green Destinations has 3 main categories we can start with:  

* Energy use
* Waste & Recycling
* Water management

Alban has mentioned they keep "social metrics" as well, so if future data surfaces than this goal could address some of the other themes listed above as well. 


## Set up 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(here)

source(here('workflow/R/common.R'))


##file paths

goal     <- 'tr'
scenario <- 'v2020'
dir_goal      <- file.path(here('prep',goal, scenario))
dir_github    <- '~/github/tet-prep'

brando_stats <- read_csv(file.path(dir_goal,"_raw_data/tr_stats_beachcomber.csv"))
```


## Methods

For this goal, we will score each sustainability category based on two criteria: 

1. How close it is to 100% sustainable, or zero impact?
2. Is this category being closely monitored?

If the answer to the second criteria is "no", the score will be reduced by 0.5.

Based on our limited information, we believe that energy and waste meet both these criteria, while water does not. Furthermore, it is unclear how much water sourced from non-native freshwater is actually being used, so its difficult to give this category a score. 

If a category has multiple metrics, those will be averaged to get the category score. We will then average the three category scores to get the final score. 

```{r}

#energy - only use the amount of total energy that is produced via renewable efforts, the air conditioning stat isn't helpful without knowing how much air conditioning makes up of all energy usage (although its probably a lot) 

energy_score <- brando_stats %>% 
  filter(category == "energy") %>% 
  filter(percent == 0.7) %>%
  mutate(
    sustainability = percent,
    monitoring = 1,
    score = percent*monitoring
  ) %>% 
  dplyr::select(category, units, score)


#waste - average the three waste scores, since each are a different type of waste

waste_score <- brando_stats %>% 
  filter(category == "waste") %>%
   mutate(
    sustainability = percent,
    monitoring = 1,
    score = percent*monitoring) %>% 
  summarize(
    score = mean(score)
  ) %>% 
  mutate(
    category = "waste",
    units = "waste management"
  )


# water - what percent of water used by the resort is sourced from rainwater, desal or other non-freshwater sources 

water_score <- brando_stats %>% 
  filter(category == "water") %>% 
  filter(percent == 1) %>% 
   mutate(
    sustainability = percent,
    monitoring = 0.5,
    score = percent*monitoring) %>% 
  dplyr::select(category, units, score = percent)

```

Bind the dfs together and find the average, creating status layer 

```{r}

tr_scores <- bind_rows(energy_score, waste_score, water_score) %>% 
  summarize(
    status = mean(score)
  ) %>% 
  mutate(
    region_id = 1,
    year = 2020
  )

```

For now, save this to the int folder (once we finalize it, save to outputs and toolbox)

```{r}

write_csv(tr_scores, file.path(dir_goal, "int/tr_status_final.csv"))


```

We were not able to find any information on The Brando website or the Beachcomber sustainability document on how the sustainability of the resort has evolved since its opening in 2014. Therefore, for now we will provide the same score for every year since 2014.

```{r}

years <- seq(2014,2020,1)

tr_status <- data.frame(
  region_id = rep(1, length(years)),
  status = rep(tr_scores$status, length(years)),
  year = years
)

#write out final layer to output and toolbox
write_csv(tr_status, file.path(dir_goal, "output/tr_status.csv"))


write_csv(tr_status, file.path( "~/github/tet-scores/region/layers/tr_status.csv"))

```

