---
title: 'OHI Tetiaroa: Lasting Special Places'
author: "Maddie"
output:
  html_document:
    highlight: haddock
    includes:
      in_header: '~/github/tet-prep/workflow/templates/tet_hdr.html' 
    number_sections: yes
    theme: cerulean
    toc: yes
---

## Summary

The methods for the global OHI assessment use the extent of areas protected as a way to measure the aesthetic, spiritual, and cultural value people hold for a certain place. The model for this goal considers the inland coastal zone (up to 1 km inland) independently from, and equally weighted with, the offshore coastal zone (up to 3 nm offshore).  The status for this goal is calculated as:  

<br>

$$X_{LSP} = \frac{\left(\frac{Area_{P}}{Area_{P_{ref}}} + \frac{Area_{MPA}}{Area_{MPA_{ref}}}\right)}{2}$$

where: 

* $Area_{P}$ = Protected area for inland 1 km buffer
* ${Area_{P_{ref}}}$ = Reference value for inland protected area
* $Area_{MPA}$ = Marine protected area for offshore 3 nm buffer
* ${Area_{MPA_{ref}}}$ = Reference value for marine protected area within offshore 3 nm buffer
* $Ref$ = 30% of total area within buffer zone is protected
  
    
<br>

Because Tetiaroa is such a small area, we will adjust these methods. First, instead of just calculating protected area within 1km of the shore, we will treat all land areas as coastal land on Tetiaroa, since most motus (individual islands) are not wider than 1 km anyway. Second, instead of creating a buffer of 3nm and treating that as all offshore, we will split up the marine area into two categories: the lagoon area, and the area extending outwards 3nm from each motu ("offshore"). 

Lastly, we will adjust what we define as "area protected", computing a score for each zone based on if they meet four criteria, developed through conversations with Frank Murphy and Neil Davies: 

1. Is there more than one stakeholder?
2. Is there a management plan for that area in place?
3. Does there exist some sort of monitoring or tracking of the effects of the management plan?
4. Has the management plan been effective in changing behaviors or meeting is objectives?

Below is a table listing these criteria for different management zones in Tetiaroa, with a 1 indicating this criteria is being fulfilled to the best extent, and a 0 indicating that a criteria does not exist at all:  

| Zone | Multiple Stakeholders? | Plan? | Monitoring? | Effective? |
| :---- | :----------- | :------- | :------- | :------- |
| All motus | 1 | 1 | 1 | 1 |
| No-take zone | 1 | 1 | 1 | 0.75 |
| Rest of Lagoon | 1 | 1 | 1 | 1 |
| 3nm offshore | 0 | 0.5 | 0 | 0 | 1 |

**Motus**

The land in Tetiaroa is privately owned by Marlon Brando's trust, which has a 99-year simple fee lease. Brando always wanted Tetiaroa to be preserved as an ecological sanctuary, and therefore helped found The Tetiaroa society to manage all conservation activities on the island. The Tetiaroa Society has worked with the trust, the government and The Brando resort to create management plans for all motus. These areas meet all four criteria.

**Lagoon No-Take Zone**. 

Half the lagoon is considered a protected "no-take zone", which is enforced by rangers employed with the Tetiaroa Society. According to the Tetiaroa Society's [sustainable use plan](https://www.tetiaroasociety.org/sites/default/files/research-docs/Tetiaroa-CASUP.pdf), this area is recognized by the Ministry for Natural Resources as of June 2014. However, Frank and Neil pointed out that while there are rangers employed, the monitoring effort has not helped to determine how well this area is deterring fisherman or other ecologically damaging activities. Therefore, it meets with first three criteria and scores a 0.75 in its effectiveness. 

**Rest of the lagoon**

The Tetiaroa Society and The Brando are still involved in the management of the rest of the lagoon, so we consider this area to have multiple stakeholders. There is also technically a plan, as seasonal fishing restrictions are in place and enforced. Frank feels this enforcement is pretty effective, more so than the no-take zone. This zone therefore meets 4 out of 4 criteria.

**Offshore**

The marine area 3 nm offshore from Tetiaroa is managed solely by the French Polynesian Government. While there exist restrictions on fishing in all FP waters (ie year-round prohibitions on shark fishing), there does not seem to exist a comprehensive management plan for pelagic ecosystems, so we'll score this criteria as 0.5. Because the existing "management plan" is very simple, Frank and Neil feel that it is being enforced well, even if it is not monitored well. This might seem counterintuitive, but it may be that fishers and visitors are adhering to the restrictions more because violation can result in a large fine and possible jail time, acting a strong deterrent despite a lack of monitoring.

The final score will be the average of the land, lagoon and offshore. We will calculate the land score by calculating the weighted average of each motu's score, which leaves room for the score of a specific motu to change if future monitoring finds that the management plans have not been effective in things like deterring poachers or preventing introductions of native species. 


***

## Setup

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      message = FALSE, warning = FALSE, eval=FALSE)

library(raster)
library(sf)
library(tidyverse)
library(here)

source(here('workflow/R/common.R'))

goal     <- 'lsp'
scenario <- 'v2020'
dir_goal      <- file.path(here('prep/sp',goal, scenario))
dir_goal_anx  <- file.path(dir_anx, goal, scenario)
dir_github    <- '~/github/tet-prep'


```

## Methods

### Check for overlapping MPAs

First, we'll check the WDPA raster and confirm there aren't any MPAs that fall within our 3nm buffer. Since these all came out as NA, we can conclude there aren't any MPAs in the offshore area. 

``` {r, check_mpas, eval = FALSE}
#source wdpa raster
rast_wdpa <- raster::raster(file.path(dir_M, 'git-annex/globalprep/lsp/v2020/rast/wdpa_2020_moll_500m.tif')) #note: the numbers for this are 2019, does that seem right?

#to convert into raster needs to be in an sp object instead sfc_POLYGON
tet_buffer_3nm_sp <- as(tet_buffer_3nm,"Spatial")

crop <- crop(rast_wdpa, extent(tet_buffer_3nm_sp), snap="out")

plot(crop)

#this comes out empty, probably not any MPAs in this region 
```

***

### Calculate region areas 

We'll calculate the total area of our buffer, and then subtract the areas of the motus and the no-take zone, which we know is 8869 acres in size. 

``` {r calc_rgn_area, eval = FALSE}

#buffered area = total area of offshore and onshore we are interested in
calc_area_3nm <- st_read(file.path(dir_github, 'spatial/shp'), layer = 'tet_buffer_3nm') %>% 
  mutate(area_m2 = st_area(.),
         area_km2 = area_m2/1000000) %>% 
  separate(area_km2, into = c("area_km2", "units"), sep = " ") %>% 
  dplyr::select(-area_m2, -units)

area_3nm <- as.numeric(calc_area_3nm$area_km2)

total_tet_area <- data.frame(
  region = "total",
  year = 2020,
  area_km2 = area_3nm
)

#export to outputs folder

write_csv(total_tet_area, file.path(dir_goal, 'int/total_tet_area.csv'), append = FALSE)
```


Finding the area of the motus and of the lagoon
``` {r calc_prot_area, eval = FALSE}
#motus - some have one stakeholder, while others have more than one.
#calculate area
area_motus <- st_read(file.path(dir_github, 'spatial/shp'), layer = 'tet_motus_area') %>% 
  dplyr::select(region = name, area_km2) %>% 
  st_set_geometry(.,NULL) %>% 
  mutate(
    year = 2020
  )

#lagoon - half of this is protected with a no take zone.
#convert acres into kilometers squared
##since the no take area is appx. half the lagoon, we can assume the rest of it is the same area.

no_take_area_km <- 0.00404686*8869

lagoon_df <- data.frame(
  region =  c("No take zone","Rest of lagoon"),
  area_km2 = rep(no_take_area_km,2),
  year = rep(2020, 2)
)

 
#create df with the motus and the lagoon, which we'll subtract from the total to get the offshore + half the lagoon 

motus_lagoon_area <- bind_rows(area_motus, lagoon_df) %>% 
  summarize(
    total_area_km = sum(area_km2)
  )

#offshore
#now subtract this total from the total to get the offshore area NOT protected, and combine that all into a df

offshore_area <- total_tet_area[1,3] - motus_lagoon_area[1,1]

all_regions_df <- data.frame(
  region = "Offshore",
  area_km2 = offshore_area,
  year = 2020
) %>% 
  bind_rows(lagoon_df, area_motus)

write_csv(all_regions_df, file.path(dir_goal, 'int/all_regions_area.csv'))

```

### Assign scores

Next step is to assign a score of 0 - 4 for each area in the data frame and divide by 4 to normalize to 0 - 1.  

We will calculate the total score by averaging across the land, lagoon and offshore scores. 

```{r}

all_regions <- read_csv(file.path(dir_goal,"int/all_regions_area.csv")) %>% 
  mutate(
    score = case_when(
      region %in% "No take zone" ~ 3.75,
      region %in% "Offshore" ~ 1.5,
      region %in% "Rest of lagoon" ~ 4,
      TRUE ~ 4
    )
  ) %>% 
  mutate(
    status = score / 4
  )


#note: these calculations will be done in the toolbox, but doing them here now so we can see what the scores would be
land_status <- all_regions %>% 
  filter(region %!in% c("Offshore","No take zone","Rest of lagoon")) %>% 
  summarize(
    status = weighted.mean(status,area_km2)
  ) #1

lagoon_status <- all_regions %>% 
  filter(region %in% c("No take zone","Rest of lagoon")) %>% 
  summarize(
    status = mean(status)
  ) #0.625

lsp_status <- all_regions %>% 
  filter(region == "Offshore") %>% 
  dplyr::select(status) %>% 
  bind_rows(land_status, lagoon_status) %>% 
  summarize(
    status = mean(status) #0.54
  ) %>% 
  mutate(
    region_id = 1,
    year = 2020
  )
```


***

## Write out layers

From the protected area file created, write out the individual layers ready for the Toolbox[TM]. These need to include years (2014 - 2020). Since they resort started construction in 2009, and the island was purchased in 1967, most of the ownership can remain the same. The no take zone was established in 2014 so that has not changed either, although if we decide to extend the trend calculation beyond 5 years than it would change.

- `lsp_status_mgmt.csv` areas of each zone and its score for years 2014 - 2020

``` {r write_layers, eval = FALSE}

years = rep(seq(2014,2020,1), 16) %>% 
  sort()


lsp_status_final <- data.frame(
    region_id = 1,
    zone = rep(all_regions$region, length(years)),
    area = rep(all_regions$area_km2, length(years)),
    status = rep(all_regions$status, length(years)),
    year = years
    )
#output
write_csv(lsp_status_final, file.path(dir_goal, "output/lsp_status_mgmt.csv"))

#toolbox

write.csv(lsp_status_final, file = "~/github/tet-scores/region/layers/lsp_status.csv", row.names = FALSE)


```

## Resilience: Management

We will also use these scores as a resilience layer, since better management will produce positive outcomes for species, habitats, tourism and the conservation economy. Essentially, this will effect all the goals as this layer affects all areas considered in the analysis. 

Since we've already created it above, we just need to write it to the resilience folder for record keeping, and to the toolbox under a different name. 

```{r}

write.csv(lsp_status_final, file = "~/github/tet-scores/region/layers/res_mgmt.csv", row.names = FALSE)

write.csv(lsp_status_final, file.path(dir_github, "/prep/resilience/social/management/res_mgmt.csv"))

```

